---
title: "Lab 14: Summarizing Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts.
Your task is to fill in the required code snippets.

## Birds Data Again

Today we will be working with the birds data once again (and as on
Thursday, hopefully allowing time to once again look at Project B):

```{r}
birds <- read_csv("https://statsmaths.github.io/stat_data/birds.csv")
```

Here is a full data dictionary describing all of the variables

- **genus** (chr): taxonomic rank of the bird
- **species** (chr): scientific species name of the bird
- **name** (chr): common name of the bird
- **type** (chr): common name of the type of bird
- **egg_mass** (dbl): average mass of an egg when laid (grams)
- **male_mass** (dbl): average observed mass of an adult male (grams)
- **mating_system** (int): scores of mating system. Intensity of male-male competition increases from 1 to 5.
    - (1) polyandry
    - (2) monogamy (<5% polygyny)
    - (3) mostly monogamy, but occasional polygyny (5–15% polygyny)
    - (4) mostly polygyny (> 15% polygyny)
    - (5) lek or promiscuous
- **display** (int): mating display agility
    - (1) ground displays only, including displays on trees and bushes
    - (2) ground displays, but with occasional jumps/leaps into the air
    - (3) both ground and non-acrobatic flight displays
    - (4) mainly aerial displays, non-acrobatic
    - (5) mainly aerial displays, acrobatic
- **resource** (int): scores of territoriality and between-mate resource sharing
    - (0) males and females don't share resources and they feed away from their breeding territory
    - (1) males and females share resources on their territory only during the breeding season
    - (2) males and females share resources on their territory all year round.
- **clutch_size** (dbl): average number of eggs produced per clutch

Notice that the last two variables are integer codes. They are stored as numbers but
correspond to a category.

## Summarizing the Data

Create a new dataset which summarizes everything to the level of a bird 
type using the `group_summarize` function:

```{r}
bird_types <- group_summarize(birds, type)
```

Create a log-log plot of the median egg mass and median male mass for each
bird type.

```{r}
ggplot(bird_types, aes(egg_mass_median, male_mass_median)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10()
```

Now create a new dataset filtered to only those bird types with 5 or more
birds in the original dataset:

```{r}
major_bird_types <- filter(bird_types, n >= 5)
```

Finally, add two layers to the first plot that (1) colors the smaller set 
of birds in red and (2) provides a textual label for the red colored points.
Try to make the labels readable (hint: perhaps add the `alpha` aesthetic
to the first set of points and/or jitter the y-axis of the labels):

```{r}
ggplot(bird_types, aes(egg_mass_median, male_mass_median)) +
  geom_point(alpha=0.15) +
  geom_point(data = major_bird_types, color="red") +
  annotate("text", x = 266, y = 3900, label="Albatross") +
  annotate("text", x = 2.99, y = 27.5, label="Bunting") +
  annotate("text", x = 1.3, y = 12, label="Fairywren") +
  annotate("text", x = 1.33, y = 6.3, label="Gerygone") +
  annotate("text", x = 3.3, y = 19, label="Honeyeater") +
  annotate("text", x = 31, y = 337.5, label="Owl") +
  annotate("text", x = 6.4, y = 71.35, label="Parrot") +
  annotate("text", x = 60, y = 442, label="Petrel") +
  annotate("text", x = 2, y = 19, label="Robin") +
  annotate("text", x = 7.5, y = 28, label="Sandpiper") +
  annotate("text", x = 2.74, y = 22.6, label="Sparrow") +
  annotate("text", x = 1.32, y = 8, label="Warbler") +
  scale_x_log10() +
  scale_y_log10()
```


